Data Analysis Methods for the Detection of the Epoch of Reionization

### 2017-11-29 PAPER.tex
some minor edits

%		Math Shortcuts from Adrian
%\def\b{\mathbf{b}}
%\def\k{\mathbf{k}}
%\def\r{\mathbf{r}}
%\def\q{\mathbf{q}}
%\def\b{\mathbf{b}}
%\def\kp{\mathbf{k}^\prime}
%\def\kpp{\mathbf{k}^{\prime\prime}}
%\def\V{\mathbb{V}}
%\def\At{\tilde{A}}
%\def\Vt{\tilde{V}}
%\def\Tt{\tilde{T}}
%\def\tb{\langle T_b\rangle}
%\newcommand{\vis}{\mathbf{v}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\C}{\mathbf{C}}
\newcommand{\Chat}{\mathbf{\hat{C}}}
\newcommand{\F}{\mathbf{F}}
\newcommand{\Fhat}{\mathbf{\hat{F}}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\I}{\mathbf{I}}
%\newcommand{\xhat}{\hat{\mathbf{x}}}
%\newcommand{\nhat}{\hat{\mathbf{n}}}
%\newcommand{\A}{\mathbf{A}}
%\newcommand{\N}{\mathbf{N}}
%\newcommand{\rhat}{\hat{\mathbf{r}}}
%\newcommand{\khat}{\hat{\mathbf{k}}}
%\newcommand{\btheta}{\boldsymbol \theta}

@@

By about one billion years after the Big Bang ($z \sim 6$), the very first stars and galaxies are thought to have ionized all the 
neutral hydrogen that dominated the baryonic matter content in the Universe. This transition period, during which the first 
luminous structures formed from gravitational collapse and began to emit intense radiation that ionized the cold neutral gas 
into a plasma, is known as the epoch of reionization (EoR). The EoR is a relatively unexplored era in our cosmic dawn. Its 
history encodes important information regarding the nature of the first galaxies and the processes of structure formation. 
Direct measurements of the EoR would unlock powerful information about the intergalactic medium, revealing connections 
between the smooth matter distribution exhibited via cosmic microwave background (CMB) studies and the highly structured 
web of galaxies we observe today.

One promising technique to probe the EoR is to target the 21 cm wavelength emission that is emitted by neutral hydrogen via 
its spin-flip transition. This technique is powerful because it can be observed as a function of redshift --- that is, the wavelength 
of the signal reaching our telescopes can be directly mapped to a distance from where the emission originated before 
stretching out as it traveled through expanding space. Tracing the 21 cm line as a function of redshift offers a window into the 
evolution of ionization, temperature, and density fluctuations during this transformative era.

Although a detection of the EoR remains elusive, there are several radio telescope experiments that have succeeded in using 
the 21 cm signal from hydrogen to place constraints on the brightness of the EoR. Examples of experiments investigating the 
mean brightness temperature of the EoR relative to the CMB are the Experiment to Detect the Global EoR Signature (EDGES; 
\citealt{bowman2010}), the Large Aperture Experiment to Detect the Dark Ages (LEDA; \citealt{greenhill_bernardi2012}), the 
Dark Ages Radio Explorer (DARE; \citealt{burns2012}), the Sonda Cosmol\'ogica de las Islas para la Detecci\'on de 
Hidr\'ogeno NeutroSciHi (SCI-HI; \citealt{voytek2014}), the Broadband Instrument for Global HydrOgen ReioNisation Signal 
(BIGHORNS; \citealt{sokolowski2015}), and the Shaped Antenna measurement of the background RAdio Spectrum (SARAS; 
\citealt{patra2015}). Radio interferometers which seek to measure statistical power spectra include the Giant Metre-wave 
Radio Telescope (GMRT; \citealt{paciga_et_al2013}), the LOw Frequency ARray (LOFAR; \citealt{van_haarlem_et_al2013}), 
the Murchison Widefield Array (MWA; \citealt{tingay_et_al2013}), the 21 Centimeter Array (21CMA; 
\citealt{peterson_et_al2004}; \citealt{wu2009}), and PAPER (\citealt{parsons_et_al2010}). The Hydrogen Epoch of 
Reionization Array (HERA), which is currently being built, is a next-generation instrument that aims to combine lessons 
learned from previous experiments and is forecasted to be able to make a successful high-significance power spectrum 
detection with an eventual $350$ elements using current analysis techniques (\citealt{deboer_et_al2017}; 
\citealt{pober_et_al2014}).

The major challenge that faces all 21 cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, 4-5 orders of magnitude brighter. A clean measurement therefore requires 
an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three themes that are essential to investigate 
for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER, motivating a revised PAPER-64 $21$ cm power spectrum 
from the lessons learned.

@@

A deep understanding of the following three themes is essential for the accuracy and interpretation of a $21$ cm power 
spectrum result. Stemming from a re-analysis of PAPER-64 data, we believe these themes serve as an important check-list for 
a rigorous power spectrum analysis.
\begin{itemize}
\item \textbf{Signal Loss} (Section \ref{sec:SiglossOverview}): Certain analysis techniques can lead to the loss of the EoR 
signal. If not corrected for, this loss could lead to a false non-detection. Computing signal loss correctly has subtle challenges 
but is necessary to ensure the accuracy of any result. 
\item \textbf{Error Estimation} (Section \ref{sec:ErrorOverview}): Confidence intervals on the $21$ cm power spectrum result 
determine the difference between a detection and a null result, which have two very different implications. Errors can be 
estimated in a variety of ways, and we will discuss a few of them.
\item \textbf{Bias} (Section \ref{sec:BiasOverview}): There are several possible sources of positive power offset in a visibility 
measurement that can show up as a detection in a power spectrum, such as bias from noise and foregrounds. In particular, a 
successful EoR detection would also imitate a bias. Proving a bias is EoR may be the most difficult challenge for $21$ cm 
analyses, as it is crucial to be able to distinguish a detection of foreground leakage, for example, from that of EoR. In this paper 
we will highlight some sources of bias, discuss ways to mitigate their effects, and describe tests that a true EoR detection must 
pass.
\end{itemize}

@@

In the next three subsections, we study each theme in depth, focusing on how power spectrum technique trade-offs affect each. 
We use a toy data model to develop intuition into why certain analysis choices may be appealing and discuss ways in which 
they are limited. We highlight problems that can arise regarding each theme and offer suggestions to mitigate the issues. 
Ultimately, we show that rigorous investigations into signal loss, error estimation, and bias must be considered for robust $21$ 
cm results.

@@

In this section we illustrate how signal loss arises in analysis. Signal loss refers to attenuation of the target cosmological signal 
in a power spectrum estimate. It can arise in a variety of ways in the analysis pipeline, such as fitting a polynomial during 
spectral calibration, applying a delay-domain filter, or by weighting data by itself. Here we focus on signal loss associated with 
applying a weighting matrix to data, a loss that can be significant depending on the choice of weighting and one that was 
previously underestimated in the P64 analysis

@@

We begin with our data matrix, $\textbf{x}$, which contains our measured visibilities in Jy. It has length ($N_{t}, N_{f}$), where 
$N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. Visibilities are measurements of 
the Fourier transform of the sky along $2$ spatial dimensions, and since we are interested in $3$-dimensional Fourier-modes 
we only need to take one Fourier transform of our data along the line-of-sight dimension. We do this when forming the un-
normalized power spectrum estimate $\hat{q}_{\alpha}$:

@@

\noindent The matrix $\textbf{Q}$ is an operator that takes our frequency-domain visibilities and Fourier-transforms them into 
power spectrum space, while also taking into account physical constants such as the Boltzmann constant used to convert Jy to 
Kelvin, as well as cosmological scalings. The index $\alpha$ denotes a waveband in $k_{\parallel}$, where $k_{\parallel}$ is the 
Fourier-dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}). $\textbf{R}$ is a weighting matrix --- as 
an example, inverse covariance weighting would set $\textbf{R} = \textbf{C}^{-1}$ and an unweighted case would use $
\textbf{R} = \textbf{I}$, the identity matrix.

@@

\noindent where $\hat{\textbf{p}}$ is the normalized estimate of the true power spectrum. The data analyst has a choice for $
\textbf{M}$ --- for simplicity in this section we set $\textbf{M} \propto \textbf{I}$, although we explore other cases for the analysis 
of PAPER-64 data as explained in Section \ref{sec:Bias}.

@@

To build our intuition into how a particular choice of weighting, namely inverse covariance weighting, can lead to signal loss, we 
use a toy model. We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time 
integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which 
might be used for a power spectrum analysis. For PAPER-64 we use $\sim$$8$ hours of data for our analysis (with channel 
widths of $0.5$ MHz and integration times of $43$ seconds), but here we scale it down for this toy model with no loss of 
generality.

We create mock visibilities, $\textbf{x}$, and assume a non-tracking, drift scan observation. Hence, flat spectrum sources (away 
from zenith) oscillate in time and frequency in our measurements. We therefore form a mock bright foreground signal, $
\textbf{x}_{FG}$, as a complex sinusoid that varies smoothly in time and frequency, a simplistic but realistic representation of a 
single bright source. We also create a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal. A more 
realistic EoR signal would have a sloped power spectrum in $\hat{\textbf{p}}$(k) (instead of flat, as in the case of white noise), 
which could be simulated by introducing frequency correlations into the mock EoR signal. However, in this paper we treat all 
$k$'s separately, so a simplistic white noise approximation can be used. Our combined data vector is then $\textbf{x} = 
\textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes. The three data vectors are shown in Figure 
\ref{fig:toy_sigloss1}.

@@

If the true covariance matrix $\textbf{C}$ of our data was known, then every single eigenvalue of $\textbf{C}$ would be 
representative of real fluctuations in the data. However, when using an estimated $\hat{\textbf{C}}$ that is derived from one 
particular data realization, its eigenvalues may differ from the truth. Said differently, shapes that may not exist (or have a 
weaker existence) in a true covariance may appear stronger in the estimated covariance. Hence, they will be down-weighted 
more than they should be.

@@

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield zero 
signal loss. Recall that our toy model foreground is a sinusoid that can be perfectly described by a single eigenmode. We can 
project out the zeroth eigenmode (i.e. zero out all eigenmodes except the first one), thereby down-weighting the foregrounds 
perfectly and nothing else. Hence, we are not using information from all the weaker eigenmodes which carry the risk of 
describing EoR fluctuations. Altering $\hat{\textbf{C}}$ as such is one example of a regularization method, in which we are 
changing $\hat{\textbf{C}}$ in a way that flattens its eigenspectrum, making it more similar to that of $\textbf{I}$. The resulting 
power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover EoR, 
demonstrating that if we can model our foregrounds perfectly (and disentangle them from other modes), we can down-weight 
them without signal loss. There are several other ways to regularize $\hat{\textbf{C}}$, and we will discuss some in Section 
\ref{sec:otherweight}.

@@

We next form $500$ bootstraps ($N_{boot} = 500$). To create each bootstrap, we draw $N$ samples, with replacement, of the 
original data, and take the mean over the $N$ samples. The standard deviation over the $500$ bootstraps gives an error 
estimate for our dataset. This error is indicated by the grey star in Figure \ref{fig:toy_error1} and matches our theoretical 
prediction (green).

@@

In Figure \ref{fig:toy_error2} we compare the two methods of bootstrapping: sampling all elements randomly (black points) 
versus sampling just the final element randomly (grey points). The two converge for datasets with small numbers of elements, 
when filling a few spots randomly is nearly the same as filling only the last spot randomly. However, there is also increased 
scatter in this regime due to small number statistics.

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},width=\columnwidth]{plots/toy_error2.png}
	\caption{The fraction of independent elements in one bootstrap sample as a function of the total number of elements in a 
dataset (i.e. if $4$ out of $5$ elements are independent, this would yield a fraction of $0.8$). A fraction of $1.0$ means that all 
elements in the bootstrap sample are independent, and therefore sensitivity is maximized (alternately, this axis can be thought 
of as power spectrum sensitivity, where sensitivity increases moving upwards along the y-axis). Two bootstrapping methods are shown here --- sampling all elements with replacement (grey) and sampling only the last element with replacement (black). The former fails to maximize sensitivity since some elements end up being repeated due to random sampling.} 
%The `outlier' points lying at a fraction of $1.0$ at the far left of the plot occur in the off-chance that the last spot in the sample happens to be the one missing element instead of a repeated element --- this is more likely to happen with a small total number of elements but also results in a more dramatically different fraction as compared to the case of a repeated element in the last spot, hence why these points appear like outliers.}
	\label{fig:toy_error2}
\end{figure}

@@

To capture the full statistical likelihood of signal loss, one requires a quick way to generate many realizations of simulated $21$ 
cm signal visibilities. Here we use the same method used in P64, where mock Gaussian noise visibilities (mock EoR signals) 
are filtered in fringe-rate space to retain only ``sky-like" time-modes. This signal is then added to the data.  \footnote{One 
specific change from P64 is that we add this simulated signal into the analysis pipeline before the final fringe-rate filter is 
applied to the data. Previously, the addition was done after that final fringe-rate filter step.  This change results in an increased 
estimate of signal loss (by a factor of $\sim$$10$), likely due to the use of the fringe-rate filter as a simulator. Whether the loss would be estimated at the same level if different $21$ cm simulations were used remains an open question.}

@@

We are interested in quantifying how much variance in $\textbf{e}$ is lost after weighting $\textbf{r}$ and estimating the power 
spectrum according to OQE formalism. We investigate this by comparing two quantities we call the input power spectrum and 
output power spectrum: $P_{in}$ and $P_{out}$, defined as (ignoring normalization factors):

\begin{equation}
\label{eq:Pin}
P_{in,\alpha} \propto \textbf{e}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e}
\end{equation}

@@

\begin{eqnarray}
\label{eq:sigloss}
P_{out,\alpha} &=& \hat{\textbf{p}}_{r,\alpha}-\hat{\textbf{p}}_{x,\alpha} \nonumber \\
&\propto& \textbf{r}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{r} - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}
^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}

$P_{in}$ represents the unweighted power spectrum of $\textbf{e}$, our simulated EoR signal, 
$P_x$ is the power spectrum of the data, $P_r$ is the power spectrum of the
data plus injection, and $P_out$ is the difference in power caused by the injected signal.

One may wonder why $P_{out}$ cannot be computed simply as the weighted power spectrum of $\textbf{e}$ alone, namely 
$P_{out,\alpha} \propto \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e}$. In short, this naive 
expression fails to include correlations between the EoR signal and data, which emerges as statistical noise that is added to the 
estimate, decreasing the value of $P_{out}$. Mathematically, we can show this by expanding out Equation \eqref{eq:sigloss}:

\begin{eqnarray}
\label{eq:crossterm}
P_{out,\alpha} &\propto& (\textbf{x}+\textbf{e})^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}(\textbf{x}+\textbf{e}) - 
\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x} \nonumber \\
&\propto& \textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} + \textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}
^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}

We see that $P_{out}$ is comprised of multiple terms. Taking the case of very large $\textbf{e}$ so that any terms involving only 
$\textbf{x}$ are small, yields:

>>>>

%		Math Shortcuts from Adrian
%\def\b{\mathbf{b}}
%\def\k{\mathbf{k}}
%\def\r{\mathbf{r}}
%\def\q{\mathbf{q}}
%\def\b{\mathbf{b}}
%\def\kp{\mathbf{k}^\prime}
%\def\kpp{\mathbf{k}^{\prime\prime}}
%\def\V{\mathbb{V}}
%\def\At{\tilde{A}}
%\def\Vt{\tilde{V}}
%\def\Tt{\tilde{T}}
%\def\tb{\langle T_b\rangle}
%\newcommand{\vis}{\mathbf{v}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\f}{\mathbf{f}}
\newcommand{\s}{\mathbf{s}}
\newcommand{\e}{\mathbf{e}}
\newcommand{\p}{\mathbf{p}}
\newcommand{\phat}{\hat{\mathbf{p}}}
\newcommand{\C}{\mathbf{C}}
\newcommand{\Chat}{\mathbf{\hat{C}}}
\newcommand{\F}{\mathbf{F}}
\newcommand{\Fhat}{\mathbf{\hat{F}}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\I}{\mathbf{I}}
%\newcommand{\xhat}{\hat{\mathbf{x}}}
%\newcommand{\nhat}{\hat{\mathbf{n}}}
%\newcommand{\A}{\mathbf{A}}
%\newcommand{\N}{\mathbf{N}}
%\newcommand{\rhat}{\hat{\mathbf{r}}}
%\newcommand{\khat}{\hat{\mathbf{k}}}
%\newcommand{\btheta}{\boldsymbol \theta}

@@

By about one billion years after the Big Bang ($z \sim 6$), the very first stars and galaxies are thought to have ionized all the 
neutral hydrogen that dominated the baryonic matter content in the Universe. This transition period, during which the first 
luminous structures formed from gravitational collapse and began to emit intense radiation that ionized the cold neutral gas 
into a plasma, is known as the epoch of reionization (EoR). The EoR is a relatively unexplored era in our cosmic dawn. Its 
history encodes important information regarding the nature of the first galaxies and the processes of structure formation. 
Direct measurements of the EoR would unlock powerful information about the intergalactic medium, revealing connections 
between the matter distribution exhibited via cosmic microwave background (CMB) studies and the highly structured 
web of galaxies we observe today.

One promising technique to probe the EoR is to target the 21 cm wavelength emission that is emitted and absorbed by neutral hydrogen via 
its spin-flip transition. This technique is powerful because it can be observed both spatially and as a function of redshift --- that is, the wavelength 
of the signal reaching our telescopes can be directly mapped to a distance from where the emission originated before 
stretching out as it traveled through expanding space. Tracing the 21 cm line as a function of redshift offers a window into the 
evolution of ionization, temperature, and density fluctuations during this transformative era.

Although a detection of the EoR remains elusive, there are several radio telescope experiments that have succeeded in using 
the 21 cm signal from hydrogen to place constraints on the brightness of the signal. Examples of experiments investigating the 
mean brightness temperature of the EoR relative to the CMB are the Experiment to Detect the Global EoR Signature (EDGES; 
\citealt{bowman2010}), the Large Aperture Experiment to Detect the Dark Ages (LEDA; \citealt{greenhill_bernardi2012}), the 
Dark Ages Radio Explorer (DARE; \citealt{burns2012}), the Sonda Cosmol\'ogica de las Islas para la Detecci\'on de 
Hidr\'ogeno NeutroSciHi (SCI-HI; \citealt{voytek2014}), the Broadband Instrument for Global HydrOgen ReioNisation Signal 
(BIGHORNS; \citealt{sokolowski2015}), and the Shaped Antenna measurement of the background RAdio Spectrum (SARAS; 
\citealt{patra2015}). Radio interferometers which seek to measure statistical power spectra include the Giant Metre-wave 
Radio Telescope (GMRT; \citealt{paciga_et_al2013}), the LOw Frequency ARray (LOFAR; \citealt{van_haarlem_et_al2013}), 
the Murchison Widefield Array (MWA; \citealt{tingay_et_al2013}), the 21 Centimeter Array (21CMA; 
\citealt{peterson_et_al2004}; \citealt{wu2009}), and PAPER (\citealt{parsons_et_al2010}). The Hydrogen Epoch of 
Reionization Array (HERA), which is currently being built, is a next-generation instrument that aims to combine lessons 
learned from previous experiments and is forecasted to be able to make a successful high-significance power spectrum 
detection with an eventual $350$ elements using current analysis techniques (\citealt{deboer_et_al2017}; 
\citealt{pober_et_al2014}; \citealt{liu_parsons_2016}).

The major challenge that faces all 21 cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter. A clean measurement therefore requires 
an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three themes that are essential to investigate 
for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER, motivating a revised PAPER-64 $21$ cm power spectrum 
from the lessons learned.

@@

A deep understanding of the following three themes is essential for the accuracy and interpretation of a $21$ cm power 
spectrum result. Stemming from a re-analysis of PAPER-64 data, we believe these themes serve as an important check-list for 
a rigorous power spectrum analysis.
\begin{itemize}
\item \textbf{Signal Loss} (Section \ref{sec:SiglossOverview}): Signal loss refers to attenuation of the target cosmological signal 
in a power spectrum estimate. Certain analysis techniques can cause this loss, and if not corrected for, it could lead to false non-detections and overly aggressive upper limits. Computing signal loss correctly has subtle challenges but is necessary to ensure the accuracy of any result. 
\item \textbf{Error Estimation} (Section \ref{sec:ErrorOverview}): Confidence intervals on the $21$ cm power spectrum result 
determine the difference between a detection and a null result, which have two very different implications. Errors can be 
estimated in a variety of ways, and we will discuss a few of them.
\item \textbf{Bias} (Section \ref{sec:BiasOverview}): There are several possible sources of positive power offset in a visibility 
measurement that can show up as a detection in a power spectrum, such as bias from noise and foregrounds. In particular, a 
successful EoR detection would also imitate a bias. Proving a bias is an EoR detection may be the most difficult challenge for $21$ cm 
analyses, as it is crucial to be able to distinguish a detection of foreground leakage, for example, from that of EoR. In this paper 
we will highlight some sources of bias, discuss ways to mitigate their effects, and describe tests that a true EoR detection must 
pass.
\end{itemize}

@@

In the next three subsections, we study each theme in depth, focusing on how power spectrum technique trade-offs affect each. 
We use a toy data model to develop intuition into why certain analysis choices may be appealing and discuss ways in which 
they are limited. We highlight problems that can arise regarding each theme and offer suggestions to mitigate the issues. 
Ultimately, we show that rigorous investigations into signal loss, error estimation, and bias must be performed for robust $21$ 
cm results.

@@

Signal loss can arise in a variety of ways in the analysis pipeline, such as fitting a polynomial during 
spectral calibration, applying a delay-domain filter, or by weighting data by itself. Here we focus on signal loss associated with 
applying a weighting matrix to data, a loss that can be significant depending on the choice of weighting and one that was 
previously underestimated in the P64 analysis

@@

We begin with our data matrix, $\textbf{x}$, which contains our measured visibilities for a single baseline in Jy. It has length ($N_{t}, N_{f}$), where 
$N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. Visibilities are measurements of 
the Fourier transform of the sky along $2$ spatial dimensions (using the flat-sky approximation), and since we are interested in $3$-dimensional Fourier-modes 
we only need to take one Fourier transform of our data along the line-of-sight dimension. We do this when forming the un-
normalized power spectrum estimate $\hat{q}_{\alpha}$:

@@

\noindent $\textbf{Q}$ is a family of matrices that takes our frequency-domain visibilities and Fourier-transforms them into 
power spectrum space, while also taking into account physical constants such as the Boltzmann constant used to convert Jy to 
Kelvin, as well as cosmological scalings. The index $\alpha$ denotes a waveband in $k_{\parallel}$, where $k_{\parallel}$ is the 
Fourier-dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}). $\textbf{R}$ is a weighting matrix --- as 
an example, inverse covariance weighting would set $\textbf{R} = \textbf{C}^{-1}$ and an unweighted case would use $
\textbf{R} = \textbf{I}$, the identity matrix.

@@

\noindent where $\hat{\textbf{p}}$ is the normalized estimate of the true power spectrum. The data analyst has a choice for $
\textbf{M}$ --- for simplicity in this section we set $\textbf{M} = \textbf{I}$, although we explore other cases for the analysis 
of PAPER-64 data as explained in Section \ref{sec:Bias}.

@@

To build our intuition into how a particular choice of weighting, namely inverse covariance weighting, can lead to signal loss, we 
use a toy model. We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time 
integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which 
might be used for a power spectrum analysis. For PAPER-64 (both the previous analysis and our new analysis)we use $\sim$eight hours of data for our analysis (with channel 
widths of $0.5$ MHz and integration times of $43$ seconds), but here we scale it down for this toy model with no loss of 
generality.

We create mock visibilities, $\textbf{x}$, and assume a non-tracking, drift scan observation. Hence, flat spectrum sources (away 
from zenith) oscillate in time and frequency in our measurements. We therefore form a mock bright foreground signal, $
\textbf{x}_{FG}$, as a complex sinusoid that varies smoothly in time and frequency, a simplistic but realistic representation of a 
single bright source. We also create a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal. A more 
realistic EoR signal would have a sloped power spectrum in $\textbf{p}$(k) (instead of flat, as in the case of white noise), 
which could be simulated by introducing frequency correlations into the mock EoR signal. However, in this paper we treat all 
$k$'s separately, so a simplistic white noise approximation can be used. Our combined data vector is then $\textbf{x} = 
\textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes. The three data vectors are shown in Figure 
\ref{fig:toy_sigloss1}.

@@

If the true covariance matrix $\textbf{C}$ of our data was known, then every single eigenvalue and eigenvector of $\textbf{C}$ would be 
representative of real fluctuations in the data. However, when using an estimated $\hat{\textbf{C}}$ that is derived from one 
particular data realization, its eigenvalues and eigevectors may differ from the truth. Said differently, shapes that may not exist (or have a 
weaker existence) in a true covariance may appear stronger in the estimated covariance. Hence, they will be down-weighted 
more than they should be.

@@

Using what we've learned about the eigenspectrum, we can tweak it in a simple way to suppress foregrounds and yield zero 
signal loss. Recall that our toy model foreground is a sinusoid that can be perfectly described by a single eigenmode. We can 
project out the zeroth eigenmode (i.e. zero out all eigenmodes except the first one), thereby down-weighting the foreground-dominating mode only. Hence, we are not using information from all the weaker eigenmodes which carry the risk of 
describing EoR fluctuations. Altering $\hat{\textbf{C}}$ as such is one example of a regularization method, in which we are 
changing $\hat{\textbf{C}}$ in a way that flattens its eigenspectrum, making it more similar to that of $\textbf{I}$. The resulting 
power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover EoR, 
demonstrating that if we can disentangle the foreground-dominating modes from EoR-dominating modes, we can down-weight 
them without signal loss. There are several other ways to regularize $\hat{\textbf{C}}$, and we will discuss some in Section 
\ref{sec:otherweight}.

@@

We next form $500$ bootstraps ($N_{boot} = 500$). To create each bootstrap, we draw $N$ samples, with replacement, of the 
original data, and take the mean over the $N$ samples. The standard deviation over the $500$ bootstraps gives an error 
estimate for our dataset. This error is indicated by the gray star in Figure \ref{fig:toy_error1} and matches our theoretical 
prediction (green).

@@

In Figure \ref{fig:toy_error2} we compare the two methods of bootstrapping: sampling all elements randomly (black points) 
versus sampling just the final element randomly (gray points). The two converge for datasets with small numbers of elements, 
when filling a few spots randomly is nearly the same as filling only the last spot randomly. However, there is also increased 
scatter in this regime due to small number statistics.

@@

\begin{figure}
	\centering
	\includegraphics[trim={0.3cm 0.3cm 0.3cm 0.3cm},width=\columnwidth]{plots/toy_error2.png}
	\caption{The fraction of independent elements in one bootstrap sample as a function of the total number of elements in a 
dataset (i.e. if $4$ out of $5$ elements are independent, this would yield a fraction of $0.8$). A fraction of $1.0$ means that all 
elements in the bootstrap sample are independent, and therefore sensitivity is maximized (alternately, this axis can be thought 
of as power spectrum sensitivity, where sensitivity increases moving upwards along the y-axis). Two bootstrapping methods are shown here --- sampling all elements with replacement (gray) and sampling only the last element with replacement (black). The former fails to maximize sensitivity since some elements end up being repeated due to random sampling.} 
%The `outlier' points lying at a fraction of $1.0$ at the far left of the plot occur in the off-chance that the last spot in the sample happens to be the one missing element instead of a repeated element --- this is more likely to happen with a small total number of elements but also results in a more dramatically different fraction as compared to the case of a repeated element in the last spot, hence why these points appear like outliers.}
	\label{fig:toy_error2}
\end{figure}

@@

To capture the full statistical likelihood of signal loss, one requires a quick way to generate many realizations of simulated $21$ 
cm signal visibilities. Here we use the same method used in P64, where mock Gaussian noise visibilities (mock EoR signals) 
are filtered in fringe-rate space to retain only ``sky-like" time-modes. This signal is then added to the data.  \footnote{One 
specific change from P64 is that we add this simulated signal into the analysis pipeline before the final fringe-rate filter is 
applied to the data. Previously, the addition was done after that final fringe-rate filter step.  This change results in an increased 
estimate of signal loss (by a factor of $\sim$$10$), likely due to the use of the fringe-rate filter as a simulator.}

@@

We are interested in quantifying how much variance in $\textbf{e}$ is lost after weighting $\textbf{r}$ and estimating the power 
spectrum according to OQE formalism. We investigate this by comparing two quantities we call the input power spectrum and 
output power spectrum: $P_{\rm in}$ and $P_{\rm out}$, defined as (ignoring normalization factors):

\begin{equation}
\label{eq:Pin}
P_{\rm in,\alpha} = \textbf{e}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e}
\end{equation}

@@

\begin{eqnarray}
\label{eq:sigloss}
P_{\rm out,\alpha} &=& \hat{\textbf{p}}_{r,\alpha}-\hat{\textbf{p}}_{x,\alpha} \nonumber \\
&=& \textbf{r}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{r} - \textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}
^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}

$P_{\rm in}$ represents the unweighted power spectrum of $\textbf{e}$, our simulated EoR signal, 
$\hat{\textbf{p}}_{x}$ is the power spectrum of the data, $\hat{\textbf{p}}_{r}$ is the power spectrum of the
data plus injection, and $P_{\rm out}$ is the difference in power between the two.

In short, we can summarize our signal loss computation as the ratio of $P_{\rm out}/P_{\rm in}$, evaluated at the data level $\textbf{x}$ (though our actual detailed computation is described in the next section). We motivate the fact that we can evaluate the output-to-input power spectrum ratio at $\textbf{x}$ by the following reasoning.

In a limit with no instrumental noise, the data that we measure, $\x$, is comprised of two signals, such that
\begin{equation}
\x \equiv \f + \s,
\end{equation}
where $\mathbf{f}$ represents the foregrounds and $\mathbf{s}$ represents the cosmological signal. In general, suppose that our power spectrum algorithm passes the data through some function $g$, yielding a lossy estimate of the power spectrum $\phat$. This can be parametrized as
\begin{equation}
\label{eq:LinearPspecSum}
\langle \phat \rangle  = \langle g(\x) \rangle = c_{\rm fg} \p_{\rm fg} + c_{\rm eor} \p_{\rm eor},
\end{equation}
where $c_{\rm eor}$ and $c_{\rm fg}$ are multiplicative factors accounting for the signal loss in the true EoR power spectrum $\p_{\rm eor}$ and true foreground power spectrum $\p_{\rm fg}$, respectively. It is not \emph{a priori} obvious why this parameterization is suitable; we thus provide a toy model in Appendix \ref{sec:sigloss_appendix} to motivate this, although it should be noted that the result is an approximation.
